---
title: "Gov 50 Final Project"
author: "Aditi Ambravan"
description: "An Analysis of Judicial Checks and Balances"
output:
  distill::distill_article:
    self_contained: false
---
## Introduction

Right from learning about the federalist papers’ vision for judicial checks and balances, the concept of a strong, independent judiciary has fascinated me. This project is an exploration of judicial checks and balances and how judicial systems impact political freedom across the world. I will be utilizing the dataset from the Dartmouth Tuck School of Business on Judicial Checks and Balances to explore the question: What aspects of a judiciary contribute to the reinforcement or weakening of political freedom? Specifically, I aim to investigate how Judicial Independence and Constitutional Review determine political freedom. Political freedom is universally defined as the ability of a nation's citizens to participate freely in the political process, and a nation’s judiciary, given that it handles free speech issues, often has a large impact on political freedom. Political freedom can be measured by a joint assessment of a country’s democracy score and index of political rights. I hypothesize that a more independent judiciary with an ability for high-level constitutional review correlates with stronger political freedom. I reason that stronger, more independent judiciaries can adequately check legislative and executive actions to determine whether or not they infringe on a citizen’s political freedom. 

The relevance of judicial independence and political freedom has become evident in recent decisions by the United States Supreme Court. As the judiciary becomes more intertwined with the executive branch due to recent court appointments, we have witnessed landmark court cases being overturned, leading to claims that such actions violate fundamental principles of checks and balances. My curiosity about the U.S. judicial system has expanded globally, prompting several questions. Can judiciaries in other countries act similarly? Do countries lacking political freedom have weak judicial systems? It is these inquiries that motivated my selection of this dataset.

## The Data

The data source I am exploring was developed by Dartmouth’s Tuck School of Business in 2004. The data includes various dependent and independent variables and represents constitutional rules in 71 countries using different variables that reflect the strength of judicial checks and balances. The countries included are those in the Maddex Encyclopedia of Constitutions, except those with transition economies (whose constitutions are rapidly changing). 

<aside>
Dataset Citation:
"'Judicial Checks and Balances,' (with F. López-de-Silanes, C. Pop-Eleches, and A. Shleifer), Journal of Political Economy, 2004."
</aside>

It is important to note that the creators of the data did not measure the dependent and independent variables set themselves but were drawn from existing measurements and assessments. Specifically, I will be focusing on four variables in my analysis. The dependent variable, political freedom, is a variable that I creeated as a combination of democracy scores and the index of political rights. Both of these variables are measured on an international scale and are drawn from publicly available data. The independent variables are constitutional review and judicial independence, which represent the overall ability of a judiciary to engage in checks and balances. The Constitutional Review value is computed by the dataset authors as Constitutional review is computed as the normalized sum of (i) the judiciary review index and (ii) the rigidity of the constitution index. The Judicial Independence Score is computed by the dataset authors as the normalized sum of (i) the tenure of Supreme Court judges (ii) the tenure of administrative Court judges (iii) the case law variable. 

My study is cross-sectional. Additionally, the data was collected at a single point in time, meaning it does not fill the criteria for a before/after or differences-in-differences study.

```{r}
## Set up Chunk, Data Cleaning, and Creation of New Variables
library(readxl)
library(tidyverse)
library(ggplot2)
jcb_data <- read_excel("~/Downloads/jcb_data.xls")
```

## The Creation of Two New Variables

The authors of the dataset indicated that political freedom is best measured by a joint assessment of the political rights index and democracy score. Therefore, I used these two provided variables to create a normalized score for political freedom. I added the Z-scores of each variable, as a normalized score should have the same mean and standard deviation to maintain the integrity of the data. The process of creating this dependent variable for political freedom can be seen below.

The second variable that I created was a region variable. With 71 countries in the dataset, I wanted to be able to group the countries by region to simplify my visualization of the dependent variable. The process of creating this region variable can be seen below. 

```{r}
## Creation of dependent variable of political freedom

# Drop NA values from data
jcb_data <- drop_na(jcb_data, pright)

# Calculate aggregate index via Z-score sum
jcb_data <- jcb_data |>
  mutate(rights_zscore = (pright - mean(pright)) / sd(pright)) |>
  mutate(democracy_zscore = (P_dem94 - mean(P_dem94) / sd(P_dem94))) |>
  mutate(dependent_variable = rights_zscore + democracy_zscore)

## Creation of Region Variable
jcb_data <- jcb_data |>
  mutate(region_master = case_when(europe == 1 ~ "Europe",
                                   africa == 1 ~ "Africa",
                                   c_amer == 1 ~ "Central America",
                                   n_amer == 1 ~ "North America",
                                   oceania == 1 ~ "Oceania",
                                   meast == 1 ~ "Middle East",
                                   america == 1 ~ "America",
                                   asia == 1 ~ "Asia",
                                   aurasia == 1 ~ "South Asia"))
```

## Summary Plot of the Dependent Variable

```{r}
dependent_variable_plot <- jcb_data |>
  ggplot(aes(x = region_master, y = dependent_variable)) +
  geom_boxplot(fill = '#FFCBCB') +
  labs(title = "Political Freedom by Region",
       x = "Region",
       y = "Political Freedom Value")
dependent_variable_plot
```
My dependent variable for this project is the political freedom variable that I created, which combines the Z-scores of a country's political rights index and democracy score. The higher the political freedom value, the more politically free a country is, meaning that individuals are free from oppression or coercion. From the plot above, we can see that Europe, North America, and Oceania have the highest median political freedom. On the other hand. Africa and the Middle East have the lowest median political freedom scores.

## Results

## Regression 1: Political Freedom and Judicial Independence
My first independent variable of interest is judicial independence. The following regression table displays the details for a regression equation involving only that variable. 

```{r}
## Regression Calculation: Political Freedom and Judicial Independence
judicial_independence_regression <- lm(dependent_variable ~ new_independence, data = jcb_data)
modelsummary::modelsummary(judicial_independence_regression,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))
```

We can see in the table above that the coefficient associated with constitutional review is 4.727, which also represents the slope. We can interpret this to mean that the political freedom score increases by 1 point each time there is an associated 4.727 point increase in the judicial independence. This number has a p-value of 0.014, which is statistically significant with an alpha of 0.05. However, due to the presence of numerous confounders and factors that play into a country's political freedom, I do not believe that this relationship can be interpreted causally.

Below is a graph representing this regression.

```{r}
## Plot of Political Freedom and Judicial Independence
judicial_regression_plot <- jcb_data |>
  ggplot(aes(x = new_independence, y = dependent_variable)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, color = "cornflowerblue") +
  labs(
    title = "Political Freedom vs. Judicial Independence",
    x = "Judicial Independence",
    y = "Political Freedom"
  )
judicial_regression_plot
```

## Regression 2: Political Freedom and Constitutional Review
My second independent variable of interest is constitutional review. The following regression table displays the details for a regression equation involving only that variable.

```{r}
## Regression Calculation: Political Freedom and Constitutional Review
constitutional_review_regression <- lm(dependent_variable ~ new_constitutional, data = jcb_data)
modelsummary::modelsummary(constitutional_review_regression,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))
```

We can see in the table above that the coefficient associated with constitutional review is 5.177, which also represents the slope. We can interpret this to mean that the political freedom score increases by 1 point each time there is an associated 5.177 increase in the constitutional review score. This number has a p-value of 0.020, which is statistically significant with an alpha of 0.05. However, due to the presence of numerous confounders and factors that play into a country's political freedom, I do not believe that this relationship can be interpreted causally.

Below is a graph representing this regression.

```{r}
## Plot of Political Freedom and Constitutional Review
constitutional_regression_plot <- jcb_data |>
  ggplot(aes(x = new_constitutional, y = dependent_variable)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, color = "darkorchid1") +
  labs(
    title = "Political Freedom vs. Constitutional Review",
    x = "Constitutional Review",
    y = "Political Freedom")
constitutional_regression_plot
```

## Multiple Regression

After conducting the two individual regressions, I am now curious to see the coefficients under a multiple regression model. Doing so allows me to assess the strength of the relationship between the dependent variable, political freedom, and multiple predictor variables once. The following table dispays the details for a multiple regression.

```{r}
## Multiple Regression for both Independent Variables
multiple_regression <- lm(dependent_variable ~ new_independence + new_constitutional, data = jcb_data)
modelsummary::modelsummary(multiple_regression,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"))
```

In the table above, the variable with a larger effect on political freedom is judicial independence, which has a coefficient of 4.152, meaning that as we hold all else equal, political freedom increases by 1 point as judicial independence increases by 4.152 points. On the other hand, the constitutional review variable has a smaller effect, with a coefficient of 3.069. This means that as we hold all else equal, political freedom increases by 1 point as constitutional review increases by 3.069 points. Thus, we can determine that greater judicial independence is a stronger determinant of political freedom than stronger constitutional review. The judicial independence coefficient has a p-value of 0.034, which implies statistical significance at the alpha level of 0.05. The constitutional review coefficient has a p-value of 0.179, which implies a lack of statistical significance at the alpha level of 0.05. It is important to note that a multiple regression cannot be visualized in a 2D format. 


## Conclusion
Ultimately, both judicial independence and constitutional review have an impact on political freedom. However, from the multiple regression, I am more confident that judicial independence has a statistically significant impact on political freedom than constitutional review because of its greater regression coefficient and p-value that is below my selected alpha of 0.05, which implies statistical significance. Based on these results, my hypothesis is half-supported, that stronger judicial independence correlates with greater political freedom. The results support the rejection of the null hypothesis that judicial independence has no effect on political freedom. However, there are limitations to the analysis due to the presence of confounding variables that are not a part of the study. There are immeasurable factors within countries that can impact political freedom, from nationalistic sentiments to political strife. Another limitation is the lack of data for some countries, where the dataset inputs an N/A value due to countries not being able to report or access certain measurements. With more time, I could improve my analysis by including more variables within the multiple regression that may influence political freedom to gain a clearer picture on what variables most impact political freedom in a statistically significant way. Additionally, expanding the analysis to include more countries would allow for a more comprehensive analysis.